A flexible and intuitive analytic expression for the metallicity-dependent cosmic star formation rate

### 2022-07-01 src/tex/ms.tex
[showyourwork] overleaf sync

It is difficult to observationally constrain the shape of \SFRDzZ (see \cite{Chruslinska2019_obs} for an extensive overview and discussion of relevant observational caveats). Even at low redshifts, the low metallicity part of the distribution is poorly constrained \citep{Chruslinska+2021}.
Traditionally the metallicity density distribution is estimated by combining a mass-metallicity relation (MZ-relation) and a galaxy stellar mass function (GSMF) \citep[see also][]{Chruslinska+2018, Chruslinska2019_effectCO, Broekgaarden+2021a}.
Another way is to extract the metallicity density fraction from cosmological simulations \citep[e.g.]{Mapelli2017, Schneider+2017}
Alternatively, one can combine an observed star formation rate, \SFRDz, like from \cite{MadauDickinson2014} or \cite{Madau+2017}, and convolve this with some belief about the shape of the metallicity density distribution, such as was was done in \cite{Neijssel+2019}.
A third option is extract this information from cosmological simulations \citep[e.g.][]{Mapelli+2017,Schneider+2017}.
In this work we provide an analytical model which we fit to the \SFRDzZ based on the TNG 100 simulation \citep{Pillepich2018, Weinberger2017}.

@@

% %
% The task we have at hand here is not a standard textbook fitting problem given the specific aims and requirements that we have that are driven by the astrophysical applications we have in mind. Qualitatively, our aims are the following. Our main aim is to find solution for the free parameters, for which the expression reproduces the overall behaviour that is observed in the cosmological simulations. For the purposes we have in mind, it will be more important to prioritise fitting the large scale trends, while we are not so interested in smaller scale fluctuations. We further prioritise getting the bulk of the cosmic star formation right and are less interested in capturing the behaviour in metallicity and redshift bins where the cosmic star formation rate is very low and thus insignificant for the overall picture. Finally, we are especially interested in capturing the asymmetry of the metallicity distribution. 
% There is not one unique procedure that will full fill these criteria.  After experimenting with different criteria we have adopted the procedure below, which we believe works well for the aims we have in mind.

@@

Different cosmological simulations produce varying \SFRDzZ, depending on the adopted assumptions about the creation and distribution of metals. For a detailed discussion of the processes that govern the creation, distribution and mixing of metals in in the TNG simulations we refer to \cite{Pakmor+2022}.
We have chosen to fit of our analytical model to the illustris simulations because I) they have been shown to reproduce many of the global properties of galaxies and their scaling relations for a representative portion of the Universe \citep[e.g.][]{FirstResTNG_Naiman2018,Torrey+2021,Genel+2018,Hemler+2021}. 
II) \cite{Briel+2021} find that the \SFRDzZ from the EAGLE \citep{Schaye+2015,Crain+2015} and TNG simulations provide the best agreement between observed and predicted cosmic rates for electromagnetic and gravitational-wave transients. And III) the simulation output is publicly available.\footnote{ \url{https://www.tng-project.org/}} 
Although in this work, we only fit our model to the IllustrisTNG100 simulation, our prescription can be easily be used to fit other observational or simulated data of the metallicity dependent star formation rate density.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\script{Plot_Mass_distributions.py}
\includegraphics[width=0.8\textwidth]{figures/Mass_distributions_all_SFRD_variations.pdf}
\caption{The primary mass distribution of merging BBH systems from isolated binary evolution for several variations in \SFRDzZ. 
The first five panels show variations of the cosmic metallicity distribution  \dpdZ (eq. \ref{eq: z log skew}, parameters listed in the first two columns of Table \ref{tab: fit params}). The bottom right panel shows variations in the magnitude of the star formation rate with redshift, i.e. \SFRDz. For the latter we vary the four fiducial parameters of \SFRDz simultaneously (last two columns of Table \ref{tab: fit params}). All panels are shown at a reference redshift of $z=0.2$, with the corresponding predicted BBH merger rate annotated in the legend. We show the power-law + peak model from \protect\cite{GWTC3_popPaper2021} in grey. Lastly we annotate the relative change in the rate at three reference masses: $10\Msun$, $25\Msun$ and $40\Msun$. Variations in \SFRDzZ have the largest impact on the high mass end of the distribution, while around $\Mbheen=10$, variations are smaller than a factor of 3.  
  \label{fig: mass dists}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary \label{sec: summary}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We present a flexible analytic expression for the metallicity-dependent star formation rate, \SFRDzZ (equations \ref{eq: total sfrd}, \ref{eq: z log skew} and \ref{eq: sfr1}). An analytical expression allows for controlled experiments of the effect of \SFRDzZ on dependent rates, such as the rate and mass distribution of merging BBHs. 
The novelty of the model presented in this work is that it adopts a skewed-lognormal for the distribution of metallicities at every redshift (\dpdZ).

>>>>

It is difficult to observationally constrain the shape of \SFRDzZ (see \cite{Chruslinska2019_obs} for an extensive overview and discussion of relevant observational caveats). Even at low redshifts, the low metallicity part of the distribution is poorly constrained \citep{Chruslinska+2021}.
Traditionally the metallicity density distribution is estimated by combining a mass-metallicity relation (MZ-relation) and a galaxy stellar mass function (GSMF) \citep[see also][]{Chruslinska+2018, Chruslinska2019_effectCO, Broekgaarden+2021a}.
Another way is to extract the metallicity density fraction from cosmological simulations \citep[e.g.]{Mapelli2017, Briel+2021}. % A third option is extract this information from cosmological simulations \citep[e.g.][]{Mapelli+2017,Schneider+2017}.
Alternatively, one can combine an observed star formation rate, \SFRDz, like from \cite{MadauDickinson2014} or \cite{Madau+2017}, and convolve this with some belief about the shape of the metallicity density distribution, such as was was done in \cite{Neijssel+2019}.

%Arguments: cosmological simulations work at high z,  but are they real? observations 
In this work we provide an analytical model which we fit to the \SFRDzZ based on the TNG 100 simulation \citep{Pillepich2018, Weinberger2017}.
% %
\todo{add motivation for this shape and paper (use some of the text below) .}
The task we have at hand here is not a standard textbook fitting problem given the specific aims and requirements that we have that are driven by the astrophysical applications we have in mind. Qualitatively, our aims are the following. Our main aim is to find solution for the free parameters, for which the expression reproduces the overall behaviour that is observed in the cosmological simulations. For the purposes we have in mind, it will be more important to prioritise fitting the large scale trends, while we are not so interested in smaller scale fluctuations. We further prioritise getting the bulk of the cosmic star formation right and are less interested in capturing the behaviour in metallicity and redshift bins where the cosmic star formation rate is very low and thus insignificant for the overall picture. Finally, we are especially interested in capturing the asymmetry of the metallicity distribution. 
There is not one unique procedure that will full fill these criteria.  After experimenting with different criteria we have adopted the procedure below, which we believe works well for the aims we have in mind.  
%%

@@

@@

Different cosmological simulations produce varying \SFRDzZ, depending on the adopted assumptions about the creation and distribution of metals. For a detailed discussion of the processes that govern the creation, distribution and mixing of metals in in the TNG simulations we refer to \cite{Pakmor+2022}.
We have chosen to fit of our analytical model to the publicly available illustris simulations\footnote{ \url{https://www.tng-project.org/}}  because I) they have been shown to reproduce many of the \todo{geef concrete list (appendix other paper) } 
% The simulations were originally calibrated against the observed total cosmic star formation rate density and the stellar mass function of galaxies \citep{Pillepich2018}. They reproduce the evolution of the sizes of galaxies with redshift \citep{Genel+2018} and with observational constraints on the mass-metallicity relation of
% galaxies up to z = 2 \citep{Torrey+2019} as well as iron abundances \citep{FirstResTNG_Naiman2018} and
% the metallicity gradients within galaxies at low redshift \citep{Hemler+2021}. These simulations have also already been used to make predictions for gravitational wave sources through pairing with  predictions for the outcomes of binary evaluation obtained with the BPASS code \citet{Briel+2021}.
global properties of galaxies and their scaling relations for a representative portion of the Universe \citep[e.g.][]{FirstResTNG_Naiman2018,Torrey+2021,Genel+2018,Hemler+2021}. 
II) \cite{Briel+2021} find that the \SFRDzZ from the EAGLE \citep{Schaye+2015,Crain+2015} and TNG simulations provide the best agreement between observed and predicted cosmic rates for electromagnetic and gravitational-wave transients. 
% And III) the simulation output is publicly available.\footnote{ \url{https://www.tng-project.org/}} 
%
Although in this work, we only fit our model to the IllustrisTNG100 simulation, our prescription can be easily be used to fit other observational or simulated data of the metallicity dependent star formation rate density.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\script{Plot_Mass_distributions.py}
\includegraphics[width=0.8\textwidth]{figures/Mass_distributions_all_SFRD_variations.pdf}
\caption{The primary mass distribution of merging BBH systems from isolated binary evolution for several variations in \SFRDzZ. 
The first five panels show variations of the cosmic metallicity distribution  \dpdZ (eq. \ref{eq: z log skew}, parameters listed in the first three columns of Table \ref{tab: fit params}). The bottom right panel shows variations in the magnitude of the star formation rate with redshift, i.e. \SFRDz. For the latter we vary the four fiducial parameters of \SFRDz simultaneously (last two columns of Table \ref{tab: fit params}). All panels are shown at a reference redshift of $z=0.2$, with the corresponding predicted BBH merger rate annotated in the legend. We show the power-law + peak model from \protect\cite{GWTC3_popPaper2021} in grey. Lastly we annotate the relative change in the rate at three reference masses: $10\Msun$, $25\Msun$ and $40\Msun$. Variations in \SFRDzZ have the largest impact on the high mass end of the distribution, while around $\Mbheen=10$, variations are smaller than a factor of 3.  
  \label{fig: mass dists}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary \label{sec: summary}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{bullet point met one liners : paragraaf }
We present a flexible analytic expression for the metallicity-dependent star formation rate, \SFRDzZ (equations \ref{eq: total sfrd}, \ref{eq: z log skew} and \ref{eq: sfr1}). An analytical expression allows for controlled experiments of the effect of \SFRDzZ on dependent rates, such as the rate and mass distribution of merging BBHs. 
The novelty of the model presented in this work is that it adopts a skewed-lognormal for the distribution of metallicities at every redshift (\dpdZ).
